Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-10 Paper.Rmd
modified

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs. log(mass) for birds, \textbf{(C)} log(BMR) vs. log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of species for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript). \label{fig_biv_raw_cop}}
\end{center}
\end{figure}

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5). Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript). \label{fig_aphid_raw_cop}}
\end{center}
\end{figure}

@@

As described in the Appendix \ref{SM-Data_detailed}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (E), respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

@@

We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig. \ref{fig_biv_raw_cop}.

@@

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Fig. \ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

@@

\newpage
```{r tab_bivar_summary, echo=F, results='markup'}

>>>>

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs. log(mass) for birds, \textbf{(C)} log(BMR) vs. log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of species for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript). Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5). These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript)\label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}

<!--
\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5). Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript). \label{fig_aphid_raw_cop}}
\end{center}
\end{figure}-->

@@

As described in the Appendix \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

@@

We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig. \ref{fig_biv_multi_raw_cop}(A-H).

@@

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Fig. \ref{fig_biv_multi_raw_cop}(A) and (B) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

@@

```{r tab_bivar_summary, echo=F, results='markup'}

### 2018-10-10 SupportingInformation.Rmd
modified

d1<-data_aphid_count$`Green spruce aphid`[[4]]
d2<-data_aphid_count$`Green spruce aphid`[[11]]
pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid abundance from location 4",ylab="Aphid abundance from location 11",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

@@

# code to plot raw data and copula plot for aphid ff @location pair (2,5)
# bestfit cop : J
d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid first-flight from location 2",ylab="Aphid first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

>>>>

d1<-data_aphid_count$`Green spruce aphid`[[4]]
d2<-data_aphid_count$`Green spruce aphid`[[11]]
pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count from location 4",ylab="Green-spruce aphid count from location 11",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

@@

# code to plot raw data and copula plot for aphid ff @location pair (2,5)
# bestfit cop : J
d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling aphid's first-flight from location 2",ylab="Leaf-plum curling aphid's first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
